Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-10-16 prop_letter.tex
prop_letter.tex \section{Gleason} Updating...

To prove the theorem, we first need to construct a QIVPM, and then
verify that no states consistent with it on all projectors. Given
an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2},\ldots,\ket{d-1}\right\} $,
let $\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$, $\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$,
and
\begin{equation}
\mathscr{I}_{0}=\left\{ \necess,\imposs,\unknown\right\} \,.\label{eq:3-value-intervals}
\end{equation}
The function~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$ defined
in Table~\ref{tab:non-Born-QIVPM} can be verified as a QIVPM straightforwardly~\citep{HOSTunpublished}.
\begin{table}
\caption{\label{tab:non-Born-QIVPM}Define a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$
on the three-dimensional Hilbert space. The projectors mapping to
the same interval are listed in the column labeled by $P$ and separated
by commas.}

@@

In other words, if the measurement intervals are too imprecise, the
results we get might contradict to each other, and exclude any possible
state from being consistent with the measurement results. To address
this problem, we are going to consider a particular family of QIVPMs
induced by quantum probability measures. When we generalized quantum
probability measures to QIVPMs, we changed the range from the unit
interval~$\left[0,1\right]$ to a set of intervals~$\mathscr{I}$.
This replacement suggests us to find a function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
which can be used to replace a real-valued probability~$\mu\left(P\right)$
by an interval-valued one~$\iota\left(\mu\left(P\right)\right)$.
In other words, compositing $\iota$ with a quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
produces an interval-valued function~$\iota\circ\mu:\events\overset{\mu}{\rightarrow}\left[0,1\right]\overset{\iota}{\rightarrow}\mathscr{I}$
which has the same domain and range as a QIVPM. Although there are
infinitely many different choice of $\iota$, the function~$\iota$
satisfies the following homomorphism-like properties are of particular
interest, because it guarantee $\iota\circ\mu$ is always a QIVPM.

\begin{definition}Given a collection of intervals $\mathscr{I}$
including $\imposs$ and $\necess$, a function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
is called a \emph{morphism}\yutsung{I may need to know the precise
conditions when $\iota$ is a morphism in the sense of category theory,
or we need to find a better name for this function.}if $\iota:\left[0,1\right]\rightarrow\mathscr{I}$
is subject to the following constraints:\begin{subequations}\label{eq:iota-constraints}
\begin{eqnarray}
 &  & \iota(0)=\imposs\,,\\
 &  & \iota(1)=\necess\,,\\
 &  & \iota\left(1-x\right)=\necess-\iota\left(x\right)\,,
\end{eqnarray}
\end{subequations}and satisfies that for any three numbers~$x_{0}$,
$x_{1}$, and $x_{2}\in\left[0,1\right]$, their sum~$x_{0}+x_{1}+x_{2}\in\left[0,1\right]$
implies
\begin{equation}
\begin{aligned} & \iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\\
\subseteq{} & \iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)\,.
\end{aligned}
\label{eq:iota-Inclusion}
\end{equation}
\end{definition}

\noindent Then, $\iota$ is a morphism precisely characterizes that
$\iota\circ\mu$ is always a IVPM no matter in classical or quantum
cases.

@@

Because the similarity between classical and quantum cases, we will
only prove the equivalence between Statement~\ref{enu:iota-subject-to}
and Statement~\ref{enu:iota-mu-QIVPM}. It is clear that $\iota$
is a morphism implies $\iota\circ\mu$ is always a QIVPM, and the
rest of the proof is to prove that if $\iota$ is not a morphism,
then $\iota\circ\mu$ might not be a QIVPM. If $\iota$ doesn't satisfy
Eq.~(\ref{eq:iota-Inclusion}), we can find three particular numbers~$x_{0}$,
$x_{1}$, and $x_{2}\in\left[0,1\right]$ such that $x_{0}+x_{1}+x_{2}\in\left[0,1\right]$,
but $\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\nsubseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)$.
Consider the state $\rho=x_{0}\proj{0}+x_{1}\proj{1}+x_{2}\proj{2}+\left(1-x_{0}-x_{1}-x_{2}\right)\proj{3}$.
Then, the function~$\iota\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}:\events\rightarrow\mathscr{I}$
fails to satisfy Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
\begin{eqnarray*}
 &  & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{1}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{2}\right)\right)\\
 & \nsubseteq & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{1}+\proj{2}\right)\right)\,,
\end{eqnarray*}
and cannot be a QIVPM. Therefore, $\iota$ is a morphism if and only
if $\iota\circ\mu$ is always a QIVPM.

>>>>

To prove the theorem, we first need to construct a QIVPM, and then
verify that no states consistent with it on all projectors. Given
an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2},\ldots,\ket{d-1}\right\} $,
let $\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$, $\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$,
and
\begin{equation}
\mathscr{I}_{0}=\left\{ \necess,\imposs,\unknown\right\} \,.\label{eq:3-value-intervals}
\end{equation}
The function~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$ defined
in Table~\ref{tab:non-Born-QIVPM} can be verified as a QIVPM straightforwardly~\citep{HOSTunpublished}.
\begin{table}
\caption{\label{tab:non-Born-QIVPM}Define a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$
on a Hilbert space of dimension~$d\ge3$. The projectors mapping
to the same interval are listed in the column labeled by $P$ and
separated by commas.}

@@

In other words, if the measurement intervals are too imprecise and
the QIVPM are defined so wildly, the results we get might contradict
to each other, and exclude any possible state from being consistent
with the measurement results. To address this problem, we are going
to consider a simple rule which can induce well-behaved QIVPMs by
quantum probability measures systematically. Recall when we generalized
quantum probability measures to QIVPMs, we changed the range from
the unit interval~$\left[0,1\right]$ to a set of intervals~$\mathscr{I}$.
This replacement suggests us to find a function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
which can be used to replace a real-valued probability~$\mu\left(P\right)$
by an interval-valued one~$\iota\left(\mu\left(P\right)\right)$.
In other words, compositing $\iota$ with a quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
produces an interval-valued function~$\iota\circ\mu:\events\overset{\mu}{\rightarrow}\left[0,1\right]\overset{\iota}{\rightarrow}\mathscr{I}$
which has the same domain and range as a QIVPM. Although there are
infinitely many different choice of $\iota$, the function~$\iota$
satisfies the following homomorphism-like properties are of particular
interest, because it guarantee $\iota\circ\mu$ is always a QIVPM.

\begin{definition}Given a collection of intervals $\mathscr{I}$
including $\imposs$ and $\necess$, a function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
is called a \emph{morphism}\yutsung{Haven't found a better name so
far, but haven't proved that it is a morphism in the sense of category
theory as well...}if $\iota:\left[0,1\right]\rightarrow\mathscr{I}$
is subject to the following constraints:\begin{subequations}\label{eq:iota-constraints}
\begin{eqnarray}
 &  & \iota(0)=\imposs\,,\\
 &  & \iota(1)=\necess\,,\\
 &  & \iota\left(1-x\right)=\necess-\iota\left(x\right)\,,
\end{eqnarray}
\end{subequations}and satisfies that for any three numbers~$x_{0}$,
$x_{1}$, and $x_{2}\in\left[0,1\right]$, their sum~$x_{0}+x_{1}+x_{2}\in\left[0,1\right]$
implies
\begin{equation}
\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\subseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)\,.\label{eq:iota-Inclusion}
\end{equation}
\end{definition}

\noindent Then, $\iota$ is a morphism precisely characterizes that
$\iota\circ\mu$ is always a IVPM no matter in classical or quantum
sense.

@@

Since the similarity between classical and quantum cases, we will
only prove the equivalence between Statement~\ref{enu:iota-subject-to}
and \ref{enu:iota-mu-QIVPM}. It is clear that $\iota$ is a morphism
implies $\iota\circ\mu$ is always a QIVPM, and the rest of the proof
is to prove that if $\iota$ is not a morphism, then $\iota\circ\mu$
might not be a QIVPM. If $\iota$ doesn't satisfy Eq.~(\ref{eq:iota-Inclusion}),
we can find three particular numbers~$x_{0}$, $x_{1}$, and $x_{2}\in\left[0,1\right]$
such that $x_{0}+x_{1}+x_{2}\in\left[0,1\right]$, but $\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\nsubseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)$.
Consider the state $\rho=x_{0}\proj{0}+x_{1}\proj{1}+x_{2}\proj{2}+\left(1-x_{0}-x_{1}-x_{2}\right)\proj{3}$.
Then, the function~$\iota\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}:\events\rightarrow\mathscr{I}$
fails to satisfy Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
\begin{eqnarray*}
 &  & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{1}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{2}\right)\right)\\
 & \nsubseteq & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{1}+\proj{2}\right)\right)\,,
\end{eqnarray*}
and cannot be a QIVPM. Therefore, $\iota$ is a morphism if and only
if $\iota\circ\mu$ is always a QIVPM.
